% ************************************************
\chapter{Biology}\label{ch:Biology} 
% ************************************************


The fundamental computational units in brain networks are
neurons\index{neuron}, electrically excitable cellular elements that
process and transmit information by a cell type dependent regime of
electrical and chemical signals. Neurons are linked through
synapses\index{synapse}, forming together an expansive, interconnected
network of different neuron types, dividing into functionally and
anatomically distinct areas. The number of neurons in the average
human brain is estimated at about 86 billion, connected by
$10^{14}-10^{15}$ synapses \parencite{Herculano2009,
  Drachman2005}. Among the different brain areas studied, the
multilayered cerebral cortex\index{cortex} stands out as a region of
particular interest with many studies analyzing its structural and
dynamical features.

The principal excitatory neuron type in cortical
networks\index{cortical network} are pyramidal cells\index{pyramidal
  cell}. Connection between those neurons these are mainly of chemical
nature, releasing. While cortical networks are considered sparse,
pyramidal cells typically receive tens of thousands excitatory and
several thousand inhibitory inputs, making up for an overall
connectivity of about 0.1 in local networks \parencite{Spruston2009}.

Such synaptic contacts are inherently asymmetric. Although spike back
propagation, action potentials. From the cell body (soma\index{soma})
of the neuron. The presynaptic rewiring is the axon. Morphology of
axon and dendrite are characteristically different.

Brain networks. They are well presented by the mathematical object of
a directed graph, which will be discussed in detail in the next
chapter.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "../dplths_document"
%%% End :
