
% ######################################################################### %
% ------------------------------------------------------------------------- %
%                         Summary and Discussion
% ------------------------------------------------------------------------- %
% ######################################################################### %

\newpage
\section{Summary and Discussion}

This chapter introduced and discussed the following network types:
\begin{enumerate} \index{network types}
  \itemsep-11pt
  \item random networks
  \item distance-dependent networks
  \item rewired anisotropic networks
  \item anisotropic networks
\end{enumerate}
While random and distance-dependent networks were formally defined as
graphs in Sections~\ref{sec:random_graph_theory} and
\ref{sec:geometric_graphs}, this chapter introduced the concept of
anisotropy. Built upon it, anisotropic graphs were defined as a
geometric graphs with a specific connectivity rule. Identifying a
parameter set reflecting local cortical networks, samples of
anisotropic networks and of the other graph types were created. For
this we extracted the distance-dependent connectivity profile in
anisotropic networks and introduced rewiring as a method to eliminate
anisotropy while keeping other connectivity parameters unaltered.
Finally, by introducing the a measure for network anisotropy, we are
able to relate the concepts of distance-dependency and anisotropy,
finding that rewiring does indeed significantly reduce anisotropy
while keeping the distance-dependent connectivity intact.

The network types discussed span a spectrum of completely random
connectivity to networks with specific connectivity
rules. Investigating in how far such a specific rule is able to
produce non-random connectivity, reflecting findings in local cortical
circuits, is the main task of Chapter~\ref{ch:structural_aspects}. For
the analysis we will heavily recruit all of the network models above
as well as their computational implementations. While rewired networks
are most closely related to anisotropic networks and provide a
reference for features directly caused by anisotropy, comparison with
distance-dependent and random networks reveals indirectly affected
features and provides insight to structural features present in
anisotropic networks not found in the standard network types.


% For this we will recruit all of the networks mentioned above,
% providing a reference to test features against. 

 
% Each of the
% network types were implemented computationally and sample graphs with
% matching parameter set were created. These sample graphs will serve as
% a reference for a structural analysis, spanning the spectrum from
% completely random connectivity to specific connectivity.

% Where rewired networks provide the minimal , distance-dependent
% provide insight into comparison with a more standard model, in many
% cases comparison of structural will give important insight.


% the claim is that distance-dependent and fully rewired are more or
% less equivalent (hints of non-equivalence found in ...) and so




%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "../dplths_document"
%%% End: 
