

\section{Discussion}\label{sec:discussion}

The structural analysis of anisotropic networks revealed connectivity
characteristically different from comparable distance-dependent or
even rewired networks. Such differences however only occur when
considering patterns in connectivity of three or more neurons, pair
connections and standard network measures remain unaffected.

Accordingly, vertex degree distributions in anisotropic networks were
found to have increased variance and skew, but as rewired and
distance-dependent networks also show this feature, this is only
indirectly caused by network anisotropy. Similarly, while
an\-iso\-tro\-pic networks display a small-world property, short path
lengths and high clustering coefficients are not due to anisotropy,
but stem from the network's distance-dependency. In fact we find that
eliminating anisotropy in the network can lead to even shorter path
lengths and higher clustering, strengthening the small-worldness of
the network. Finally, we were able to present conclusive evidence that
anisotropy is not able to reproduce overrepresentation of reciprocally
connected neuron pairs, reported repeatedly in cortical networks.

In terms of higher order connectivity however, anisotropy yields
highly interesting results. Three-neuron motif occurrences are not
only strongly influenced by anisotropy, but resulting
overrepresentations resemble in many aspects the profiles found by
\textcite{Song2005}. Furthermore, anisotropy in connectivity induces
an increased frequency of high edge counts in neuron clusters,
matching results from experiments in the rat's
cortex \parencite{Perin2011}. In their study, Perin et al.\ were able
to relate such edge count overrepresentations to a \textit{common
  neighbor rule} - a specific relationship between shared in- or
outputs of a neuron pair and the probability of connection in the
pair. Being able to reproduce increased high edge counts by imposing
such relation distance-dependent networks, the common neighbor rule
was identified as an underlying organization principle. By the
interpretation of the authors, stereotypically connected clusters
arising from this rule then constitute \enquote{building blocks} of
cortical networks to be molded by experience.

In this study, anisotropy was recruited as a connection principle for
the generation of geometric networks, reflecting the dimensions of
typical cortical slices. Capturing stereotypical axonal and dendritic
morphology of pyramidal cells in the cortex, those anisotropic
networks inherently feature connectivity going beyond
distance-dependency. Indeed, not only do anisotropic networks display
non-random patterns as found in local cortical circuits, but we were
here able to show that a common neighbor rule itself is highly
affected by anisotropy. Together, these findings promote the concept
of anisotropy as an important underlying connection principle.

The results presented here provide new insight in how non-random
connectivity in cortical circuits may arise from morphological
characteristics inherent to cells in cortical networks. Employing an
abstract model, capturing stereotypical axonal and dendritic
morphology, allowed for an analytic and numerical analysis, revealing
the strong impact of anisotropy on structural aspects in neural
networks. Certainly such an oversimplified model brings caveats; as a
consequence of the chosen axon shapes, for example, directions of
axonal projections had to be chosen at random, while statistics from
the cortex imply alignment of axons. Also, the choice of positioning
of the neurons on a surface risks the loss of information contained in
three-dimensional, spatial networks. However, basing the analysis on
comparisons with otherwise identical networks lacking anisotropy, the
chosen abstractions serve the aim of this report in identifying the
role of anisotropy in determining structural aspects of geometric
neural networks as precisely as possible.

With anisotropy likely being only one amongst many potential
morpho\-logy-induced connectivity principles, the results here make an
important case for the study and inclusion of morphological aspects in
future of networks models. The presented connectivity principle may
provide a first step towards a network archetype better resembling
non-random connectivity features found in cortical circuits. As such
patterns om connectivity were found to strongly affect network
dynamics \parencite{Pernice2011, Zhao2011}, development in this
direction is crucial. Furthermore, underlying principles may reveal
unexpected, not yet reported connectivity statistics. Here we found
that the distribution of the number of common inputs for a random
neuron pair shows a drastically higher variance in networks with a
high degree of anisotropy, potentially enabling an increased
functional specificity in the networks.

As continuously evolving experimental techniques and efforts provide
more and more information on brain connectivity, identification of
connection principles will become ever more important. This work
provides an important aspect in the discussion and identification of
such underlying principles, implying that neuronal morphology may play
a crucial role in network connectivity. Countering arguments focusing
heavily on plasticity as connection principles, the observations here
may contribute to a differentiated discussion of underlying principles
including various aspects of brain networks, in the hopes of finding
consistent models for connectivity in the different areas of the
brain.

%  Insight in such statistics may
% then eventually 

% functional synaptic plasticity providing a likely
% important framework. The results presented here thus may help shift

% The results here contribute , shig

% Plastic

% Thus, important , and shifting a one-sided discussion to
% an differentiated in aspects.



% The findings call for more in which generic
% morphology might drastically influence connectivity. Such pre-existing
% restraints and may be essential in the in shaping plastic learning
% processes, Here, we provided an novel





% In this study
% find we found that anisotropy affect a common neighbor rule
% itself. Therefore. promoting to a candidate.


% Pointing out the The results are to be interpreted critically
% however, as an analytical consideration reveals 

% Following an analytical consideration of motif
% distributions in distance-dependent networks, we point out an
% important aspect in the interpretation of such distributions, finding
% an unexpected dependency


% Here, we find that not only edge count overrepresentation are by
% anisotropy, but we further find a distinct common neighbor rule in the
% , . 

 
% In their study \textcite{Perin2011} were able to relate increased
% edge-counts in neuron clusters to a \textit{common neighbor rule} - a
% specific relationship between common in- or outputs in a neuron pair
% and the probability for the pair to be connected - identifying an
% underlying principle for the occurrence of non-random connectivity in
% local circuits. Perin et al.\ suggest that (fire together wire
% together). In this structural analysis we found non-random
% connectivity resembling the findings of Song et al./ and , purely from
% a g. Identifying a common neighbor rule as the inherent, that
% non-random connectivity statistics might arrise from morphological
% restrictions rather than plastic processes.




% Taking distance-dependent networks as a reference, Perin et al.\
% were able to show network connectivity features in cortical circuits
% going beyond distance-dependency. A network model hoping to reflect
% such features, clearly is required to implement connectivity going
% beyond distance-dependent connection probabilities. With an


% Thus, anisotropy in connectivity seems . Initially reporting ,i
%  Perin
% et al.\ identify connectivity features beyond distance . Only a
% network model implementing connectivity rules going beyond
% distance-dependency can possible repo. 

% Introducing such a concept motivated
% from neuronal morphology, we here anisotropy in connectivity
% as an underlying connectivity principle 


% We find that anisotropy in connectivity induces increased high edge
% counts in neuron clusters, matching overrepresentations identified in
% local cortical circuits. 

% Comparing however to distance-dependent as opposed to rewired
% networks, we find that overrepresentation is stronger
% (\autoref{fig:perin_rew_dist}).







%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "../dplths_document"
%%% End: 